{\bf Landslide: Systematic Dynamic Race Detection in Kernel-space}

### 2012-05-09 defence/defence.tex
rearrange comments
add outline slides and a few minor edits

\newcommand\linegap{\vspace{0.2in}}
\newcommand\breakslide[1]{\begin{frame}{} \begin{center} #1 \end{center} \end{frame}}
\newcommand\related[1]{\textsuperscript{\em #1}}

% slide: outline
\begin{frame}{Outline}
	\textbf{Motivation: Concurrency debugging}
	\begin{itemize}
		\item Systematic exploration versus stress testing
		\item Challenges of kernel-space
	\end{itemize}
	\linegap

@@

	{\bf Systematic Exploration} % \related{Flanagan '05}
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Run {\em every single} interleaving
		\item Provide better debugging information
	\end{itemize}
\end{frame}

@@

% slide: outline
\begin{frame}{Outline}
\end{frame}

@@

% slide: pebbles, 15-410, simics
\begin{frame}{Environment}
	\textbf{Simics:} a full-system x86 simulator
	\begin{itemize}
		\item Students in 15-410 use Simics to test their kernels.
		\item Landslide is a transparently-embedded Simics module.
	\end{itemize}
\end{frame}

% slide: visualisation (here address challenge #1)
\begin{frame}{Anatomy of Landslide}
	\begin{center}
	\includegraphics[width=0.7\textwidth]{landslide.png}
	\end{center}
\end{frame}

% slide: using -- annotations & in-landslide instrumentation
\begin{frame}{Using Landslide}
	\begin{center}
	\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	User must tell Landslide when certain kernel events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap
	Some complicated conditions require in-Landslide instrumentation.
	\begin{itemize}
		\item Scheduler design details
	\end{itemize}
\end{frame}

\begin{frame}{Decision Trace}
	\includegraphics[width=\textwidth]{found_a_bug.png}
\end{frame}

@@

\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}

@@

	{\bf Solution:} Only consider ``relevant'' memory conflicts.
	\begin{itemize}
		\item Ignore scheduler, global objects ``known to be correct''
		\begin{itemize}
			\item Sacrifice ability to test these to efficiently test everything else.
		\end{itemize}
	\end{itemize}
\end{frame}
% slide: identifying bugs
\begin{frame}{Identifying Bugs}
	\textbf{How do we know we've found a bug?}
	\linegap

@@

% slide: outline
\begin{frame}{Outline}
\end{frame}

@@

	Race conditions
	\begin{itemize}
		\item Too many waiters allowed
		\begin{itemize}
			\item One parent thread might block forever
		\end{itemize}
		\item Accidental return to an exited thread
		% Also talk about their vanish/vanish bug
	\end{itemize}
	Deterministic bugs
	\begin{itemize}
		\item Inappropriate idling
		\begin{itemize}
			\item Waiting for a timer tick with no useful work to do
		\end{itemize}
		\item Use-after-free
		\begin{itemize}
			\item Required memory tracking to detect
		\end{itemize}
	\end{itemize}
\end{frame}

@@

	``I have made something smarter than me.''
	% Heh, if it were really smarter than me, I wouldn't have had to debug it.
	\begin{itemize}
		\item \texttt{double\_wait} - two parent threads \texttt{wait()} on one exiting child
		\item Expected: One thread reaps child, the other returns failure
		\item Got: \texttt{failed assertion `nobe->cond\_queue\_next == NULL'}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Outline}
\end{frame}

@@

\begin{frame}{Decision Points} % TODO: duplicate slide with later
	A {\bf decision point} is\ldots

@@

% TODO more of these with screenshots

@@

% TODO: run through an example

>>>>

\newcommand\linegap{\vspace{0.2in}}
\newcommand\breakslide[1]{\begin{frame}{} \begin{center} #1 \end{center} \end{frame}}
\newcommand\related[1]{\textsuperscript{\em [#1]}}

\begin{frame}{Outline}
	\textbf{Motivation: Concurrency debugging}
	\begin{itemize}
		\item Systematic exploration versus stress testing
		\item Challenges of kernel-space
	\end{itemize}
	\linegap

@@

	{\bf Systematic Exploration} \related{Simsa '10}
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Run {\em every single} interleaving
		\item Provide better debugging information
	\end{itemize}
\end{frame}

@@

\newcommand\done[1]{\textcolor{gray}{\em \small #1}}

\begin{frame}{Outline}
	\done{Motivation: Concurrency debugging}
	\begin{itemize}
		\item \done{Systematic exploration versus stress testing}
		\item \done{Challenges of kernel-space}
	\end{itemize}
	\linegap

	{\bf Tool: Landslide}
	\begin{itemize}
		\item {\bf Design and interface}
		\item {\bf Addressing challenges}
	\end{itemize}
	\linegap

	Evaluation: Finding Races
	\begin{itemize}
		\item Student user study
		\item Case study
	\end{itemize}
	\linegap

	Future Work and Conclusion
\end{frame}

@@

\begin{frame}{Environment}
	\textbf{Simics:} a full-system x86 simulator
	\begin{itemize}
		\item Students in 15-410 use Simics to test their kernels.
		\item Landslide is a transparently-embedded Simics module.
	\end{itemize}
\end{frame}

\begin{frame}{Anatomy of Landslide}
	\begin{center}
	\includegraphics[width=0.7\textwidth]{landslide.png}
	\end{center}
	% Remember to talk about solving the thread scheduling challenge
\end{frame}

\subsection{Interface}

\begin{frame}{Using Landslide}
	\begin{center}
	\includegraphics[width=0.9\textwidth]{tell_landslide.png}
	\end{center}
	User must tell Landslide when certain kernel events happen:
	\begin{itemize}
		\item When do threads become runnable / descheduled?
		\item When does the scheduler switch threads?
	\end{itemize}
	\linegap
	Some complicated conditions require in-Landslide instrumentation.
	\begin{itemize}
		\item Scheduler design details
	\end{itemize}
\end{frame}

\begin{frame}{Decision Trace}
	% TODO make the stack trace coloured
	\includegraphics[width=\textwidth]{found_a_bug.png}
\end{frame}

@@

\begin{frame}{State Space Reduction}
	State spaces grow exponentially.
	\begin{itemize}
		\item Fortunately, some sequences result in identical states.
		\item Dynamic Partial Order Reduction\related{Flanagan '05}
		\begin{itemize}
			\item Requires ``memory independence relation'' between transitions.
		\end{itemize}
	\end{itemize}
\end{frame}
\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}

@@

	{\bf Solution:} Only consider ``relevant'' memory conflicts.
	\begin{itemize}
		\item Ignore scheduler, global objects ``known to be correct''
		\begin{itemize}
			\item Sacrifice ability to test these to efficiently test everything else.
		\end{itemize}
		\item User must specify what to ignore.
	\end{itemize}
\end{frame}

\begin{frame}{Identifying Bugs}
	\textbf{How do we know we've found a bug?}
	\linegap

@@

\begin{frame}{Outline}
	\done{Motivation: Concurrency debugging}
	\begin{itemize}
		\item \done{Systematic exploration versus stress testing}
		\item \done{Challenges of kernel-space}
	\end{itemize}
	\linegap

	\done{Tool: Landslide}
	\begin{itemize}
		\item \done{Design and interface}
		\item \done{Addressing challenges}
	\end{itemize}
	\linegap

	{\bf Evaluation: Finding Races}
	\begin{itemize}
		\item {\bf Student user study}
		\item {\bf Case study}
	\end{itemize}
	\linegap

	Future Work and Conclusion
\end{frame}

@@

	Race conditions
	\begin{itemize}
		\item Too many waiters allowed
		\begin{itemize}
			\item One parent thread might block forever
		\end{itemize}
		\item Accidental return to an exited thread
		% Also talk about their vanish/vanish bug
	\end{itemize}
	\linegap

	Deterministic bugs
	\begin{itemize}
		\item Inappropriate idling
		\begin{itemize}
			\item Waiting for a timer tick with no useful work to do
		\end{itemize}
		\item Use-after-free
		\begin{itemize}
			\item Required memory tracking to detect
		\end{itemize}
	\end{itemize}
\end{frame}

@@

	``I have made something smarter than me.''
	% Heh, if it were really smarter than me, I wouldn't have had to debug it.
	\begin{itemize}
		\item \texttt{double\_wait} - two parent threads \texttt{wait()} on one exiting child
		\begin{itemize}
			\item Expected: One thread reaps child, the other returns failure
			\item Got: \texttt{failed assertion `nobe->cond\_queue\_next == NULL'}
		\end{itemize}
		\item Previously unknown, not found by stress testing or by TA
	\end{itemize}
\end{frame}

@@

\begin{frame}{Outline}
	\done{Motivation: Concurrency debugging}
	\begin{itemize}
		\item \done{Systematic exploration versus stress testing}
		\item \done{Challenges of kernel-space}
	\end{itemize}
	\linegap

	\done{Tool: Landslide}
	\begin{itemize}
		\item \done{Design and interface}
		\item \done{Addressing challenges}
	\end{itemize}
	\linegap

	\done{Evaluation: Finding Races}
	\begin{itemize}
		\item \done{Student user study}
		\item \done{Case study}
	\end{itemize}
	\linegap

@@

\begin{frame}{Decision Points} % : duplicate slide with later
	A {\bf decision point} is\ldots

@@

@@
